---
title: "QuestionB"
author: "Saeed Rahman"
date: "May 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1)

Please identify and select a specic sector's SPDR ETF, for example XLF,
XLE, etc. This site http://finviz.com/ presents details about the tick-
ers in each sector. In the sector you choose download 20 equity prices
since 2012. Perform a principal component analysis for these 20 equities.
Report the number of equities needed to account for 80% of the variability.
Using the specic PCA's you chose analyze which equities in
uence the
most these PCA's.

### Getting the data
```{r}
library(quantmod)
getSymbols("XLY",from="2012-01-01")

stockData <- new.env()
lookup.symb=c("AMZN","HD","CMCSA","DIS","MCD","PCLN","SBUX","TWX","LOW","NKE","CHTR","NFLX","TJX","GM","F","MAR","TGT","FOXA","CCL","ROST","NWL")

getSymbols(lookup.symb, from="2012-01-01", env=stockData, src="yahoo")


ReturnMatrix=NULL


for(i in 1:length(lookup.symb))
{
  tmp <- get(lookup.symb[i], pos=stockData)   # get data from stockData environment  
  ReturnMatrix=cbind(ReturnMatrix,   (Cl(tmp)-Op(tmp)) / Op(tmp)   )
  colnames(ReturnMatrix)[i]=lookup.symb[i]
}

# cor(ReturnMatrix)
```

### PCA
The output of the function PCA() is a list including :
```{r}

library("FactoMineR")

res.pca <- PCA(ReturnMatrix, graph = FALSE)
print(res.pca)
```
Variances of the principal components:
The proportion of variation retained by the principal components (PCs) can be extracted as follow :
```{r}
eigenvalues <- res.pca$eig
# head(eigenvalues[, 1:2])
head(eigenvalues)
```
Number of components that best describe 80% of variance
```{r}
eigenvalues[eigenvalues["cumulative percentage of variance"]>80,]
```
```{r}
Prin.Comp = prcomp(ReturnMatrix, scale = T) #by default R centers the variables. Scale also makes then sd=1
summary(Prin.Comp)
propvariance = Prin.Comp$sdev^2/sum(Prin.Comp$sdev^2)
plot(cumsum(propvariance), type="b")
number_best_pca= which(cumsum(propvariance)>.8)[1]
print(number_best_pca)
```
The importance of PCs can be visualized using a scree plot :
```{r}
library("factoextra")
fviz_screeplot(res.pca, ncp=10)
```
Plotting the correlations/loadings of the variables with the components
The correlation between a variable and a PC is called loading. The variables can be plotted as points in the component space using their loadings as coordinates.
```{r}
# Coordinates of variables
head(res.pca$var$coord)
fviz_pca_var(res.pca)
```

The squared loadings for variables are 
```{r}
head(res.pca$var$cos2)
```

The sum of the cos2 for variables on the principal components is equal to one.

If a variable is perfectly represented by only two components, the sum of the cos2 is equal to one. In this case the variables will be positioned on the circle of correlations.

For some of the variables, more than 2 components are required to perfectly represent the data. In this case the variables are positioned inside the circle of correlations.

```{r}
fviz_pca_var(res.pca, col.var="cos2") +
scale_color_gradient2(low="white", mid="blue", 
                    high="red", midpoint=0.5) + theme_minimal()
```





Here we can see that no variables are perfectly represented by less than two components

### Contributions of the variables to the principal components:
Variables that are correlated with PC1 and PC2 are the most important in explaining the variability in the data set.

Variables that does not correlated with any PC or correlated with the last dimensions are variables with low contribution and might be removed to simplify the overall analysis.

#### Contributions of variables on PC1
```{r}
fviz_pca_contrib(res.pca, choice = "var", axes = 1)
```

#### Contributions of variables on PC2
```{r}
fviz_pca_contrib(res.pca, choice = "var", axes = 2)
```

### Total contribution on PC1 and PC2
```{r}
fviz_pca_contrib(res.pca, choice = "var", axes = 1:2)

```

Now for further analysis in the next questions, we will select TJX, DIS, HD and ROST as the top 4 securities.

## 2)
Next pick 4 equities that are among the most in
uential in the sector you
selected in part 1. Please only select equities that trade at $5 or above.
Now we will identify a stochastic model that best ts the data for each of
them. We consider the following candidate models for each of them:

### Getting Data
```{r}
getSymbols("TJX",from="2012-01-01")
getSymbols("DIS",from="2012-01-01")
getSymbols("HD",from="2012-01-01")
getSymbols("ROST",from="2012-01-01")
stock1=ts(TJX[,6])
stock2=ts(DIS[,6])
stock3=ts(HD[,6])
stock4=ts(ROST[,6])
```

### Defining the SDE's
```{r, warning=FALSE}
library(Sim.DiffProc)
library(Ecdat)

fx <-{}
gx <-{}
#model 1 drift and diffusion 
fx[1] <- expression( theta[1]*x )
gx[1]<- expression( theta[2]*x )

#model 2 drift and diffusion 
fx[2] <- expression( theta[1]+theta[2]*x )
gx[2]<- expression( theta[3]*(x^theta[4]) )

#model 3 drift and diffusion 
fx[3] <- expression( theta[1]*x )
gx[3]<- expression (theta[2] + ( theta[3]*(x^theta[4])) )

#model 4 drift and diffusion 
fx[4]<- expression( theta[1]*x)
gx[4] <- expression( theta[2]*(x^(3/2)))

#model 5 drift and diffusion 
fx[5] <- expression( theta[1]+theta[2]*x )
gx[5] <- expression(theta[3] + (theta[4]*log(x)) )

pmle=eval(formals(fitsde.default)$pmle)

```
### Defining the functions to fit the data to the SDE's and the Maximum Likelyhood functions to estimate parameters
```{r}
Best.fit<-function(data,pmle)
{
  #model1
  mod1 <- fitsde(data=data,drift=fx[1],diffusion=gx[1],start = 
                   list(theta1=1, theta2=1,theta3=1),pmle=pmle)
  #model 2
  mod2 <- fitsde(data=data,drift=fx[2],diffusion=gx[2],start = 
                   list(theta1=1, theta2=1,theta3=1,theta4=1),pmle=pmle)
  #model 3
  mod3 <- fitsde(data=data,drift=fx[3],diffusion=gx[3],start = 
                   list(theta1=1, theta2=1,theta3=1),pmle=pmle)
  #model 4
  mod4 <- fitsde(data=data,drift=fx[4],diffusion=gx[4],start = 
                   list(theta1=1, theta2=1,theta3=1),pmle=pmle)
  
  #model 5
  mod5 <- fitsde(data=data,drift=fx[5],diffusion=gx[5],start = 
                   list(theta1=1, theta2=1,theta3=1, theta4=1),pmle=pmle)
  #Computes AIC
  AIC <- c(AIC(mod1),AIC(mod2),AIC(mod3),AIC(mod4),AIC(mod5))
  Test <- data.frame(AIC,row.names = c("Model 1","Model 2","Model 3", "Model 4","Model 5"))
  Test
  # Bestmod <- rownames(Test)[which.min(Test[,1])]
  Bestmod <- which.min(Test[,1])
  list('best.model'=Bestmod,'AIC.results'=Test)
}


Diff.mle <-function(fx,gx,data)
{
  pmle <- eval(formals(fitsde.default)$pmle)
  fitres <- lapply(1:4, function(i) fitsde(data=data,drift=fx,diffusion=gx,pmle=pmle[i],
                                           start = list(theta1=1,theta2=1,theta3=1,theta4=1)))
  Coef <- data.frame(do.call("cbind",lapply(1:4,function(i) coef(fitres[[i]]))))
  Info <- data.frame(do.call("rbind",lapply(1:4,function(i) AIC(fitres[[i]]))),
                     row.names=pmle)
  names(Coef) <- c(pmle)
  names(Info) <- c("AIC")
  list("Info"=Info,"Coef"=Coef)
}


```
### Fitting the SDE's and estimating the parameters
```{r,warning=FALSE}

pmle_type=2
best_model={}
params={}

### For Stock 1
fit1=Best.fit(data =stock1,pmle = pmle[pmle_type])
print(paste("Best model = model ",fit1$best.model))
best_model[1]=fit1$best.model

print("The parameter estimates are:")
ls1=Diff.mle(fx=fx[fit1$best.model],gx=gx[fit1$best.model],data = stock1)
ls1=ls1$Coef[,pmle_type]
params[1]=list(ls1)
print(ls1)

### For Stock 2

fit2=Best.fit(data =stock2,pmle = pmle[2])
print(paste("Best model = model ",fit2$best.model))
best_model[2]=fit2$best.model

print("The parameter estimates are:")
ls2=Diff.mle(fx=fx[fit2$best.model],gx=gx[fit2$best.model],data = stock2)
ls2=ls2$Coef[,pmle_type]
params[2]=list(ls2)


print(ls2)

### For Stock 3

fit3=Best.fit(data =stock3,pmle = pmle[pmle_type])
print(paste("Best model = model ",fit3$best.model))
best_model[3]=fit3$best.model


print("The parameter estimates are:")
ls3=Diff.mle(fx=fx[fit3$best.model],gx=gx[fit3$best.model],data = stock3)
ls3=ls3$Coef[,pmle_type]
params[3]=list(ls3)
print(ls3)

### For Stock 4

fit4=Best.fit(data =stock4,pmle = pmle[pmle_type])
print(paste("Best model = model ",fit4$best.model))
best_model[4]=fit4$best.model


print("The parameter estimates are:")
ls4=Diff.mle(fx=fx[fit4$best.model],gx=gx[fit4$best.model],data = stock4)
ls4=ls4$Coef[,pmle_type]
params[4]=list(ls4)
print(ls4)
```
### Estimating Parameters using Yuima Package
```{r}
library(yuima)
pmle_type=2
best_model={}
params={}




model1=setModel(drift="theta1*x",diffusion="theta2*x", state.var="x",time.var="t",solve.var="x",xinit=0.5)
model2=setModel(drift="theta1+theta2*x",diffusion="theta3*(x^theta4)", state.var="x",time.var="t",solve.var="x",xinit=0.5)
model3=setModel(drift="theta1*x",diffusion="theta2+theta3*(x^theta4)", state.var="x",time.var="t",solve.var="x",xinit=0.5)
model4=setModel(drift="theta1*x",diffusion="theta2*(x^(3/2))", state.var="x",time.var="t",solve.var="x",xinit=0.5)
model5=setModel(drift="theta1+theta2*x",diffusion="theta3+(theta4*log(x))", state.var="x",time.var="t",solve.var="x",xinit=0.5)


# for stock3 using yuima
print("The parameter estimates are:")

D=setYuima(data=setData(as.double(DIS[,6])),model = model3)
initialise=list(theta1=0,theta2=0,theta3=0,theta4=0)
lowbound=list(theta1=-5,theta2=-5,theta3=-5,theta4=-5)
upbound=list(theta1=5,theta2=5,theta3=5,theta4=5)
mle=qmle(D,start=initialise,lower=lowbound,upper=upbound)
summary(mle)

X=simulate(model3,true.param=list(theta1=0.140447119,theta2=-0.007829953,theta3=1.000000000,theta4=1.000000000),nsim = 1000,xinit = 39)
plot(X)

### For Stock 1
fit1=Best.fit(data =stock1,pmle = pmle[pmle_type])
print(paste("Best model = model ",fit1$best.model))
best_model[1]=fit1$best.model

m1=setModel(drift="theta*x",diffusion="sigma*x", state.var="x",time.var="t",solve.var="x",xinit=0.5)

D=setYuima(data=setData(as.double(TWX[,6])),model = m1)
print(ls1)

### For Stock 2

fit2=Best.fit(data =stock2,pmle = pmle[2])
print(paste("Best model = model ",fit2$best.model))
best_model[2]=fit2$best.model

print("The parameter estimates are:")
ls2=Diff.mle(fx=fx[fit2$best.model],gx=gx[fit2$best.model],data = stock2)
ls2=ls2$Coef[,pmle_type]
params[2]=list(ls2)


print(ls2)

### For Stock 3

fit3=Best.fit(data =stock3,pmle = pmle[pmle_type])
print(paste("Best model = model ",fit3$best.model))
best_model[3]=fit3$best.model


print("The parameter estimates are:")
ls3=Diff.mle(fx=fx[fit3$best.model],gx=gx[fit3$best.model],data = stock3)
ls3=ls3$Coef[,pmle_type]
params[3]=list(ls3)
print(ls3)

### For Stock 4

fit4=Best.fit(data =stock4,pmle = pmle[pmle_type])
print(paste("Best model = model ",fit4$best.model))
best_model[4]=fit4$best.model


print("The parameter estimates are:")
ls4=Diff.mle(fx=fx[fit4$best.model],gx=gx[fit4$best.model],data = stock4)
ls4=ls4$Coef[,pmle_type]
params[4]=list(ls4)
print(ls4)

```


## 3)
Estimate the correlation matrix  for the 4 stocks based on historical
data.
```{r}

cor(ReturnMatrix)
```
## 4)

Using the correlation matrix in the part 3 and the model you selected in
part 2 generate Monte-Carlo simulations for the 4 stocks. Even though
the Brownian motions are not exactly correlated with use
this matrix to generate correlated increments. Use T = 1 year, t = 1
255
and n = 1000 paths in your simulation. Once it is done, you only need to
keep the end value ST for each stock. You should have 4 vectors in total
each with n replications. Please report basic statistics (mean, standard
deviation, skewness, and kurtosis) for each stock.

```{r, warning=FALSE}
chol_upper=chol(cor_matrix)

n_iterations=10
n_steps=252
stocks_sim=matrix(0,n_iterations,4)
stocks_sim[,1]=stock1[1]
stocks_sim[,2]=stock2[1]
stocks_sim[,3]=stock3[1]
stocks_sim[,4]=stock4[1]

sim1={}

dt=1/n_steps

for(i in 1:n_iterations)
{
  print(i)
  for(j in 1:n_steps)
  {
    w=as.vector(matrix( rnorm(1*4,mean=0,sd=1), 1, 4))
    cor_w=chol_upper%*%w
    for(k in 1:4)
    {
     if(best_model[k]==1)
      {
        stocks_sim[i,k]=stocks_sim[i,k]+(params[[k]][1]*dt*stocks_sim[i,k])+(params[[k]][2]*stocks_sim[i,k]*w[1])
        if(i==1 && k==3 && length(sim1)<300){
          sim1=append(sim1,stocks_sim[i,k])
          print(paste("i=",i,"j=",j,"k=",k,"Price=",stocks_sim[i,k]))
          print(length(sim1))
        }
      }
      else if(best_model[k]==2)
      {
        stocks_sim[i,k]=stocks_sim[i,k]+(params[[k]][1]+params[[k]][2]*stocks_sim[i,k])*dt+
          (params[[k]][3]*stocks_sim[i,k]^params[[k]][4]*w[i])
      }
      else if(best_model[k]==3)
      {
        stocks_sim[i,k]=stocks_sim[i,k]+(params[[k]][1]*stocks_sim[i,k]*dt)+(params[[k]][2]+(params[[k]][3]*stocks_sim[i,k]^params[[k]][4]*w[i]))
      }
      else if(best_model[k]==4)
      {
        stocks_sim[i,k]=stocks_sim[i,k]+(params[[k]][1]*stocks_sim[i,k]*dt)+(params[[k]][2]*stocks_sim[i,k]^(3/2)*w[i])
      }
      else if(best_model[k]==5)
      {
        stocks_sim[i,k]=stocks_sim[i,k]+(params[[k]][1]+params[[k]][2])*dt+(params[[k]][3]+params[[k]][4]*log(stocks_sim[i,k]))*stocks_sim[i,k]*w[i]
      }
    }
    
  }
}
stocks_sim

plot(sim1)

```

















