---
title: "621 Assignment1"
author: "Saeed Rahman"
date: "February 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QUESTION 2

First function is given as:

```{r}
f1<-function(x){
  if(x==0)
    return(1)
  else
    return(sin(x)/x)
}
```
### A)
Implement the trapezoidal and the Simpson's quadrature rules to numerically
compute the denite integral above. The algorithms are presented
in [3], please see Chapter 5. Hint: you can approximate the indenite
integral by considering a large interval [????a;+a] (for example a = 106).
Consider equidistant nodes fxngN n=0; i.e., xn = ????a+n2a
N ; n = 0; 1; : : : ;N,
where N is a large integer.

### Trapezoid Rule Implementation
```{r}
TrapezoidRule = function(a, b, m, f)
{
  h = (b-a)/(m-1)
  Sum = 0.5 * h * (f(a)+f(b))
  for (i in 1:(m-2)) {
    ai = a + i*h
    Sum = Sum + h*f(ai)
  }
  return(Sum)
}

print(TrapezoidRule(b=10^6, a=-(10^6), m=12000,f1))
```
### Simpsons Rule Implementation
```{r}
SimpsonRule= function(a, b, m, f)
{
  h = (b-a)/(m-1)
  Sum = (1/3) * h * (f(a)+f(b)) 
  for (i in 1:(m-2)) {
    if(i%%2 != 0){
      Sum = Sum + h*f(a + i*h) *(4/3)
    }else{
      Sum = Sum + h*f(a + i*h) *(2/3)
    }
  }
  return(Sum)
}
print(SimpsonRule(a=-(10^6),b=10^6, m=12000,f1))
```
### B)
Compute the truncation error for the numerical algorithms implemented
in a) for a particular a 2 R and N 2 N. That is create a function of a
and N that will output IN ???? , where IN;a is the numerical approximation
of the integral. Study the changes in the approximation as N and a
increase as well as the dierence between the two quadrature approximations.
Please write your observations.
```{r}
Trapezoidal_Integration<-function(a,b,n,f)
{
  table = matrix(0, nrow=n, ncol=4, dimnames=list(
    c(1:n), c('m', 'Integration Result', 'Error', 'Error Ratio')))
  for(i in 1:n) {
    table[i,1] = ifelse(1 == i, yes=2, no=2*table[i-1,'m']-1)
    table[i,2] = TrapezoidRule(a, b, table[i,'m'], f) 
    table[i,3] = abs(table[i,2] - pi )
    table[i,4] = ifelse(1 == i, NA, table[i-1,3] /table[i,3] )
  }
  return(table)
}
Trapezoidal_Integration(a=-(10^6),b=10^6, n=24,f1)
```
```{r}
Simpson_Integration<-function(a,b,n,f)
{
  table = matrix(0, nrow=n, ncol=4, dimnames=list(
    c(1:n), c('m', 'Integration Result', 'Error', 'Error Ratio')))
  for(i in 1:n) {
    table[i,1] = ifelse(1 == i, yes=2, no=2*table[i-1,'m']-1)
    table[i,2] = SimpsonRule(a, b, table[i,'m'], f1)
    table[i,3] = abs(table[i,2]- pi )
    table[i,4] = ifelse(1 == i, NA, table[i-1,3] /table[i,3] )
  }
  return(table)
}

Simpson_Integration(a=-(10^5),b=10^5, n=20,f1)

```
From observing the convergence characterstics, it seems like that the Simpsons method get conveged at lower values of n compared to trapezoidal

## C)
In a typical scenario we do not know the true value of the integral. Thus,
to ensure the convergence of the numerical algorithms we pick a small
tolerance value " and we check at every iteration k = 1; 2; : : : if the
following condition holds:
jIk ???? Ik????1j < ";
where Ik is the value of the integral at step k. When the condition holds,
the algorithm stops. Evaluate the number of steps until the algorithms
from a) reach convergence for " = 10????4. What do you observe?

```{r}

Trapezoidal_Integration_tol<-function(a,b,n,f,tolerance=(10^-4))
{
  table = matrix(0, nrow=n, ncol=4, dimnames=list(
    c(1:n), c('m', 'Integration Result', 'Error', 'Error Ratio')))
  for(i in 1:n) {
    table[i,1] = ifelse(1 == i, yes=2, no=2*table[i-1,'m']-1)
    table[i,2] = TrapezoidRule(a, b, table[i,'m'], f) 
    table[i,3] = abs(table[i,2]-pi) 
    table[i,4] = ifelse(1 == i, NA, table[i-1,3] /table[i,3] )
    if(table[i,3] < tolerance  && table[i,1]>2){
      print(table[i,])
      print(paste("Number of Iterations=",i))
      break
  }
  }
}
```
```{r}
#Simpson_Integration_tol<-function(a,b,n,f,tolerance=(10^-4))
# {
#   table = matrix(0, nrow=n, ncol=4, dimnames=list(
#     c(1:n), c('m', 'Integration Result', 'Error', 'Error Ratio')))
#   for(i in 1:n) {
#     table[i,1] = ifelse(1 == i, yes=2, no=2*table[i-1,'m']-1)
#     table[i,2] = SimpsonRule(a, b, table[i,'m'], f1)
#     table[i,3] = abs(table[i,2]- pi)  
#     table[i,4] = ifelse(1 == i, NA, table[i-1,3] /table[i,3] )
#     if(table[i,3]<tolerance && table[i,1]>2)
#       return(table[i,])
#   }
#   # return(tail(table,1))
# 
Simpson_Integration_tol<-function(a,b,n,f,tolerance=(10^-4))
{
  table = matrix(0, nrow=n, ncol=4, dimnames=list(
    c(1:n), c('m', 'Integration Result', 'Error', 'Error Ratio')))
  for(i in 1:n) {
    table[i,1] = ifelse(1 == i, yes=2, no=2*table[i-1,'m']-1)
    table[i,2] = SimpsonRule(a, b, table[i,'m'], f1)
    table[i,3] = abs(table[i,2]- pi) 
    table[i,4] = ifelse(1 == i, NA, table[i-1,3] /table[i,3] )
    if(table[i,3] < tolerance  && table[i,1]>2){
      print(table[i,])
      print(paste("Number of Iterations=",i))
      break
    }
      
  }
}
```

```{r}
Trapezoidal_Integration_tol(a=-(10^6),b=10^6, n=24,f1) 
Simpson_Integration_tol(a=-(10^6),b=10^6, n=24,f1)
```

From the above analysis we can observe that if the range of integration is large, then Trapezoidal rule is better than Simpsons, whele if it is a small range then Simpsons converge faster in minimum number of steps

## D)


```{r}
f2<-function(x){
  return(1+(exp(-x)*sin(8*(x^(2/3)))))
}
```

```{r}

Trapezoidal_Integration_f2<-function(a,b,n,f,tolerance=(10^-4))
{
  table = matrix(0, nrow=n, ncol=4, dimnames=list(
    c(1:n), c('m', 'Integration Result', 'Error', 'Error Ratio')))
  for(i in 1:n) {
    table[i,1] = ifelse(1 == i, yes=2, no=2*table[i-1,'m']-1)
    table[i,2] = TrapezoidRule(a, b, table[i,'m'], f) 
    table[i,3] = abs(table[i,2]-2)
    table[i,4] = ifelse(1 == i, NA, table[i-1,3] /table[i,3] )
    if(table[i,3] < tolerance){
      print(table[i,])
      print(paste("Number of Iterations=",i))
      break
    }
  }
}

Simpson_Integration_f2<-function(a,b,n,f,tolerance=(10^-4))
{
  table = matrix(0, nrow=n, ncol=4, dimnames=list(
    c(1:n), c('m', 'Integration Result', 'Error', 'Error Ratio')))
  for(i in 1:n) {
    table[i,1] = ifelse(1 == i, yes=2, no=2*table[i-1,'m']-1)
    table[i,2] = SimpsonRule(a, b, table[i,'m'], f1)
    table[i,3] = abs(table[i,2]- 2)
    table[i,4] = ifelse(1 == i, NA, table[i-1,3] /table[i,3] )
    if(table[i,3] < tolerance){
      print(table[i,])
      print(paste("Number of Iterations=",i))
      break
    }
  }
}

```

```{r}

TrapezoidRule(a=0,b=2, m=20000,f2)
SimpsonRule(a=0,b=2, m=20000,f2)
```