---
title: "Question1"
author: "Saeed Rahman"
date: "March 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##A)
Construct code to calculate option values using an additive binomial tree.For this part you need for versions European and American as well as Call and Put. You may use the same tree construction for all options.

```{r}
BinomialTree = function(isCall, isAmerican=FALSE, K=100, Tm=1, 
                        S0=100, r=0.06, sig=0.2, N=3, u=1.1, d=1/u,div=0,show=FALSE)
{
  dt = Tm/N
  nu=r-div-0.5*sig*sig
  dxu=sqrt(sig*sig*dt+((nu*dt)^2))
  dxd=-dxu
  pu=0.5+0.5*(nu*dt/dxu)
  pd=1-pu
  disc=exp(-r*dt)
  #
  # dpu=disc*pu
  # dpd=disc*pd
  # edxud=exp(dxu-dxd)
  # edxd=exp(dxd)

  # Precompute constants ----
  dt = Tm/N # diving the time into partition, to get time for each state
  disc = exp(-r*dt)
  p = (exp(r*dt)-d)/(u-d) #calculating the probability of up move
  nRows = 2*N+1 #number of rows for the matrix
  nCols = N+1 #number of columns
  cp = ifelse(isCall, 1, -1) # to check ifts a call or a put
  
  # Intialize asset prices  ----
  # Creating a matrix of nRows*nColumns with zeros and headings 
  V = S = matrix(0, nrow=nRows, ncol=nCols, dimnames=list(
    paste("NumUps", N:-N, sep="="), paste("T", 0:N, sep="=")))
  S[nCols, 1] = S0 # initial stock price
  
  # iterating the elements of the matrix in a conical manner starting
  # from the position of initial stock price
  # For n=3, S[i,j]= S0, then update the forward diagonal elements 
  # Code is similar to the one used for trinomial tree
  for (j in 1:N) {
    for(i in (nCols-j+1):(nCols+j-1)) {
      S[i-1, j+1] = S[i, j]*exp(dxu)
      S[i+1, j+1] = S[i, j] *exp(dxd)
    }
  }
  for (i in 1:nRows) {
    V[i, nCols] = max( 0, cp * (S[i, nCols]-K))
  }
  # Step backwards through the tree ----
  for (j in (nCols-1):1) {
    for(i in (nCols-j+1):(nCols+j-1)) {
      # V[i, j] = disc * (p*V[i-1,j+1] + (1-p)*V[i+1,j+1])
      V[i, j] = disc * (pu*V[i-1,j+1] + pd*V[i+1,j+1])
      if(isAmerican) {
        # if american option, then take the Value at each node as the max of the
        # value of option or the payoff at that period
        V[i, j] = max(V[i, j], cp * (S[i, j] - K))
      }
    } 
  }
  if(show)
  {
    print("Stock Tree")
    print(S)
    print("Option Value Tree")
    print(V)
  }
  else
    return(V[nCols,1])
}
#To display the matrix of Stock process and Option Value
BinomialTree(isCall=TRUE,K=100,Tm =1 ,S0 =100 ,sig = .2,N =4,r=.06,show = TRUE)
print(paste("European Call Price",round(BinomialTree(isCall=TRUE,K=100,Tm =1 ,S0 =100 ,sig = .2,N=200,r=.06),4)))
print(paste("European Put Price",round(BinomialTree(isCall=FALSE,K=100,Tm =1 ,S0 =100 ,sig = .2,N =200,r=.06),4)))
print(paste("American Call Price",round(BinomialTree(isCall=TRUE,isAmerican = TRUE,K=100,Tm =1 ,S0 =100 ,sig = .2,N=200,r=.06),4)))
print(paste("American Put Price",round(BinomialTree(isCall=FALSE,isAmerican = TRUE,K=100,Tm =1 ,S0 =100 ,sig = .2,N =200,r=.06),4)))
```

##B)
Download Option prices (you can use the Bloomberg Terminal, Yahoo! Fi-
nance, etc.) for an equity, for 3 different maturities (1 month, 2 months, and
3 months) and 20 strike prices close to the value at the money. If 3 months
does not exist use next one available. Please download the data DURING
THE TRADING DAY (9:00am to 4:30pm ET). Otherwise your values will
be way off. Do not forget to download the value of the underlying. For
each strike price in the data, use the implied vol values in Homework 1 (see
Problem 1c) and the current short-term interest rate (today the Feds Fund
rate is r = 0.75%). Calculate the option price (European Calls and Puts)
using the binomial tree, and compare the results with the Black Scholes
price. Use at least 200 steps in your tree construction. Treat the options as
American as well and plot these values side by side with the European and
Black Scholes values. When you create the plot do not forget to plot the
bid-ask values as well

### Calculating Implied Volatility
```{r pressure, echo=FALSE}

library(quantmod)
#Black sholes merton pricing function----
BSM<-function(S, K, t, r, sigma,type){
  d1 <- (log(S/K)+(r+sigma^2/2)*t)/(sigma*sqrt(t))
  d2 <- d1 - sigma * sqrt(t)
  if (type == "c")
    result <- S*pnorm(d1) - K*exp(-r*t)*pnorm(d2)
  if (type == "p")
    result <- K*exp(-r*t) * pnorm(-d2) - S*pnorm(-d1)
  return(result)
}
#Calculating Implied Vol----
#Implementation of secant method that was used in assignment 1, Used secant method for the advantage in speed 
Secant <- function(S, K, t, r, type, option_price
                             , x0=0.1, x1=3, tolerance=1e-07, max.iter=10000){
  x1=3
  theta=.00001
  fun.x1=BSM(S=S,K=K,t=t,r=r,sigma=x1,type=type)-option_price
  count=1
  start.time <- Sys.time()
  while(abs(fun.x1) > tolerance && count<max.iter) {
    x2=x1-theta
    fun.x1=BSM(S=S,K=K,t=t,r=r,sigma=x1,type=type)-option_price
    fun.x2=BSM(S=S,K=K,t=t,r=r,sigma=x2,type=type)-option_price
    x1 <- x1- fun.x1/((fun.x1-fun.x2)/theta)   
    count <-count+1
  }
  end.time <- Sys.time()
  time.taken <- end.time - start.time
  if(x2<0 || count>=max.iter)
    return(list(NA,time.taken,count))
  else
    return(list(x2,time.taken,count))
}

#Calculating IV on the option chain using secant method
ImpliedVol_Secant<-function(symbol="AAPL",option_chain,rate=.75/100){
  symbol="AAPL"
  stock_df<-as.data.frame(getSymbols(symbol,from = as.Date("2017-01-01"),to=as.Date("2017-02-19"), env = NULL))

  iv <- {}
  original_iv <-{}
  optionName <-{}
  strike <-{}
  days_till_expiry <-{}
  time.taken <- 0
  iterations <- 0
  type <-{}
  bid<-{}
  ask<-{}
  for (i in 1:nrow(option_chain)) 
  {
    try({
      #Myoldmethod----
       secant <- Secant(
       S = as.numeric(tail(stock_df,1)[6]),
       K = as.numeric(option_chain[i,"Strike"]),
       t = as.numeric(option_chain[i,"days_till_expiry"])/252,
       r = rate,
       type = ifelse((option_chain[i,"Type"]=="Call"), "c", "p"),
       option_price = as.numeric(option_chain[i,"premium"]))

       iv <- append(iv,as.numeric(secant[1]))

       if(!is.na(secant[1])){
        time.taken <- as.numeric(secant[2])+time.taken
        iterations <- as.numeric(secant[3])+iterations
       }
      type <- append(type,as.character(option_chain[i,"Type"]))
      
      strike<-append(strike,as.numeric(option_chain[i,"Strike"]))
      
      optionName <- append(optionName,paste(option_chain[i,"Strike"],"-",
                                            option_chain[i,"Type"],"Expiring On:",
                                            option_chain[i,"Expiry"]))
      days_till_expiry <- append(days_till_expiry,as.numeric(option_chain[i,"days_till_expiry"]))
      
      bid<-append(bid,as.numeric(option_chain[i,"Bid"]))
      ask<-append(ask,as.numeric(option_chain[i,"Ask"]))
    })
  }
  option_chain_df <- data.frame(days_till_expiry,type,optionName,iv,strike,bid,ask)
  names(option_chain_df)<-c("Days_till_Expiry","Type","Specification","Implied_Volatility","Strike","Bid","Ask")
  time.taken <- time.taken/as.numeric(colSums(!is.na(option_chain_df))[3])
  iterations <- iterations/as.numeric(colSums(!is.na(option_chain_df))[3])
  list(option_chain_df,time.taken,iterations)
}
```

### Importing the data from the CSV files and calculating the IV 
```{r, message=FALSE, warning=FALSE}
option_chain_call_csv <- read.csv(file="call.csv",header=TRUE, sep=",")
option_chain_put_csv <-read.csv(file="put.csv",header=TRUE, sep=",")
#call.csv and put.csv are available in the project folder
#Call
option_chain_call_csv$days_till_expiry <- as.Date(option_chain_call_csv$Expiry,"%m/%d/%Y")-as.Date("2017-02-19")
#Put
option_chain_put_csv$days_till_expiry <- as.Date(option_chain_put_csv$Expiry,"%Y/%m/%d")-as.Date("2017-02-19")
#Calculating the days till expiry
option_chain_call_csv$premium<-(option_chain_call_csv$Bid+option_chain_call_csv$Ask)/2

option_chain_put_csv$premium<-(option_chain_put_csv$Bid+option_chain_put_csv$Ask)/2

df=ImpliedVol_Secant(option_chain = option_chain_call_csv)
df=as.data.frame(df[1])
options.df_call=df[complete.cases(df$Implied_Volatility),]

df=ImpliedVol_Secant(option_chain = option_chain_put_csv)
df=as.data.frame(df[1])
options.df_put=df[complete.cases(df$Implied_Volatility),]

```

### Calculating the Option Prices Using Binomial Tree and Black Sholes Merton Formula

```{r}
PriceCalculator<-function(options.df,show=FALSE){
  eur_binomial <-{}
  eur_bsm <-{}
  ame_binomial <-{}
  stock_df<-as.data.frame(getSymbols("AAPL",from = as.Date("2017-01-01"),
                                     to=as.Date("2017-02-19"), env = NULL))
  for (i in 1:nrow(options.df)){
    eur_binomial <- append(eur_binomial,BinomialTree(
                       isCall = as.logical(options.df[i,"Type"]=="Call"),
                           K  = as.numeric(options.df[i,"Strike"]),
                           Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252,
                           S0 = as.numeric(tail(stock_df,1)[6]),
                          sig = as.numeric(options.df[i,"Implied_Volatility"]),
                           r  = .75/100,
                            N = 200))
    ame_binomial <- append(ame_binomial,BinomialTree(
                    isAmerican = TRUE, 
                        isCall = as.logical(options.df[i,"Type"]=="Call"),
                            K  = as.numeric(options.df[i,"Strike"]),
                            Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252,
                            S0 = as.numeric(tail(stock_df,1)[6]),
                           sig = as.numeric(options.df[i,"Implied_Volatility"]),
                            r  = .75/100,
                             N = 200))
    eur_bsm <- append(eur_bsm,BSM(
                      type = ifelse(options.df[i,"Type"]=="Call",'c','p'),
                        K  = as.numeric(options.df[i,"Strike"]),
                        t  = as.numeric(options.df[i,"Days_till_Expiry"])/252,
                        S  = as.numeric(tail(stock_df,1)[6]),
                     sigma = as.numeric(options.df[i,"Implied_Volatility"]),
                        r  = .75/100))
                           
  }
  options.df$BinomialEuropean <- eur_binomial
  options.df$BinomialAmerican <- ame_binomial
  options.df$BlackSholesMertonEuropean <- eur_bsm
  if(show)
    head(options.df)
  else
    return(options.df)
}

PriceCalculator(options.df_call,show=TRUE)
option_chain_call <-PriceCalculator(options.df_call)
option_chain_put <-PriceCalculator(options.df_put)

```

### Plotting the Call Option Prices that are expiring in 5 days
Zoomin to see the difference in prices.
Also select legend to deactivate and activate individual series of option prices
``` {r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA, results='asis'}
# {r,echo=FALSE,results='asis',comment=NA, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

library(rCharts)
library(reshape)

option_chain_call$Implied_Volatility<-NULL
df_split <-split(option_chain_call,option_chain_call$Days_till_Expiry)
first <-df_split[1]
first <- first$`5`
first[1:2] <- list(NULL)
rownames(first) <- first$Strike
first$Strike <- NULL

a <- Highcharts$new()
a$chart(type = "line")
a$chart(zoomType="xy")
a$title(text = "Apple Call Expiring in 5 Days")
a$xAxis(categories = rownames(first),title = list(text = "Strike"),replace = T)
a$yAxis(title = list(text = "Option Price"))
# a$xAxis(title = "Strikes")
a$data(first)
a$print(include_assets = TRUE)


```

### Plotting the Call Option Prices Expiring in 26 days

```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA, results='asis'}
# {r,echo=FALSE,results='asis',comment=NA, echo=FALSE, message=FALSE, warning=FALSE, comment=NA, results='asis'}

df_split <-split(option_chain_call,option_chain_call$Days_till_Expiry)
first <-df_split[2]
first <- first$`26`
first[1:2] <- list(NULL)
rownames(first) <- first$Strike
first$Strike <- NULL

a <- Highcharts$new()
a$chart(type = "line")
a$chart(zoomType="xy")
a$title(text = "Apple Call Expiring in 26 Days")
a$xAxis(categories = rownames(first),title = list(text = "Strike"),replace = T)
a$yAxis(title = list(text = "Option Price"))
# a$xAxis(title = "Strikes")
a$data(first)
a$print(include_assets = TRUE)


```

### Plotting the Option Prices for the Call Expiring in 152 days

```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA, results='asis'}

df_split <-split(option_chain_call,option_chain_call$Days_till_Expiry)
first <-df_split[3]
first <- first$`152`
first[1:2] <- list(NULL)
rownames(first) <- first$Strike
first$Strike <- NULL

a <- Highcharts$new()
a$chart(type = "line")
a$chart(zoomType="xy")
a$title(text = "Apple Call Expiring in 152 Days")
a$xAxis(categories = rownames(first),title = list(text = "Strike"),replace = T)
a$yAxis(title = list(text = "Option Price"))
# a$xAxis(title = "Strikes")
a$data(first)
a$print(include_assets = TRUE)


```

### Plotting the Put Options Prices that are expiring in 5 days
Zoomin to see the difference in prices.
Also select legend to deactivate and activate individual series of option prices
```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA, results='asis'}
option_chain_put$Implied_Volatility<-NULL
df_split <-split(option_chain_put,option_chain_put$Days_till_Expiry)
first <-df_split[1]
first <- first$`5`
first[1:2] <- list(NULL)
rownames(first) <- first$Strike
first$Strike <- NULL

a <- Highcharts$new()
a$chart(type = "line")
a$chart(zoomType="xy")
a$title(text = "Apple Put Options Expiring in 5 Days")
a$xAxis(categories = rownames(first),title = list(text = "Strike"),replace = T)
a$yAxis(title = list(text = "Option Price"))
a$data(first)
a$print(include_assets = TRUE)

```

### Plotting the Option Prices for the Puts Expiring in 26 days

```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA, results='asis'}
df_split <-split(option_chain_put,option_chain_put$Days_till_Expiry)
first <-df_split[2]
first <- first$`26`
first[1:2] <- list(NULL)
rownames(first) <- first$Strike
first$Strike <- NULL

a <- Highcharts$new()
a$chart(type = "line")
a$chart(zoomType="xy")
a$title(text = "Apple Put Options Expiring in 26 Days")
a$xAxis(categories = rownames(first),title = list(text = "Strike"),replace = T)
a$yAxis(title = list(text = "Option Price"))
a$data(first)
a$print(include_assets = TRUE)


```

### Plotting the Option Prices for the Puts Expiring in 152 days

```{r, echo=TRUE, message=FALSE, warning=FALSE, comment=NA, results='asis'}

df_split <-split(option_chain_put,option_chain_put$Days_till_Expiry)
first <-df_split[3]
first <- first$`152`
first[1:2] <- list(NULL)
rownames(first) <- first$Strike
first$Strike <- NULL

a <- Highcharts$new()
a$chart(type = "line")
a$chart(zoomType="xy")
a$title(text = "Apple Put Options Expiring in 152 Days")
a$xAxis(categories = rownames(first),title = list(text = "Strike"),replace = T)
a$yAxis(title = list(text = "Option Price"))
a$data(first)
a$print(include_assets = TRUE)


```


##C)

The table obtained above in the previous question clearly shows that even though there exist a small error in the option prices obtained, the binomial tree(simulation method) closely predicts the price as per the Analytical equation, that is the Black sholes equation. We can also validate the accuracy of the Binomial tree and Blacksholes model used in this program by observing the small error in price difference between the calculated and the original bid and ask values.

##D)
```{r}
QuestionD<-function(){
  bsm_price <- {}
  binomial_price <-{}
  error <-{}
  iter <-c(10, 20, 30, 40, 50, 100, 150, 200, 250, 300, 350, 400)
  for(i in iter){
    binomial_temp <-BinomialTree(isCall=FALSE,K=100,Tm =1 ,S0 =100 ,sig = .2,N =i,r=.06)
    bsm_temp <-BSM(S=100,K=100,t=1,r=.06,sigma=.2,type="p")
    binomial_price <- append(binomial_price,binomial_temp)
    bsm_price <- append(bsm_price,bsm_temp)
    error <- append(error,bsm_temp-binomial_temp)
  }
  print(paste("Error=",tail(error,n=1)))
  plot(iter,error,xlab = "Number of Iterations",ylab = "Error", type="o",col="blue")
}
QuestionD()

```

It is very clear that that as the number of steps in the tree increases, the error comes down. And the error is lowest at the highest number of steps which is 400


