---
title: "Question1"
author: "Saeed Rahman"
date: "May 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A.

Implement a simple Monte Carlo scheme using m simulation trials for Eu-
ropean Call and Put options. This should be a function of n (number of
time steps) and m. In all practical applications you should use at least 300
time steps and at least 1 million simulated paths. Furthermore, implement
a calculation of the standard error of the estimate of the option price and a
way to time the simulation routine.

```{r cars}
Monte_Carlo_Pricing<-function(PutCall,S0,K,T,sigma,div,r,N,M){
  ## Precompute constants
  dt = T/N
  nudt = (r-div-0.5*sigma^2)*dt
  sigmasdt = sigma*sqrt(dt)
  lns = log(S0)
  
  sum_CT = 0
  sum_CT2 = 0
  start.time = proc.time()
  
  for (j in 1:M){
    lnSt = lns
    for ( i in 1:N){
      w = rnorm(N)
      lnSt = lnSt + nudt + sigmasdt*w
    }
    St = exp(lnSt[N])
    optionValue = ifelse(PutCall=='C', pmax(0,St-K), pmax(0,K-St))
    sum_CT = sum_CT+optionValue
    sum_CT2 = sum_CT2+optionValue*optionValue
  }
  option_value = sum_CT/M*exp(-r*T)
  SD = sqrt((sum_CT2 - sum_CT*sum_CT/M)*exp(-2*r*T)/(M-1))
  SE = SD/sqrt(M)
  end.time = proc.time()
  timetaken = end.time - start.time
  list(option_value=option_value, SE=SE, time=timetaken[3])
}
```

```{r, warning=FALSE}
# 1 Million simulations is taking long time, so for the ease of demonstration, simulations limited to 10000.
num.steps=300
num.sim=1000
r=.06
sigma=.2
div=.03
S0=100
K=100
T=1
MC_result<-Monte_Carlo_Pricing('C',S0,K,T,sigma,div,r,num.steps,num.sim)
print(paste("Price=",round(MC_result$option_value,2)))
print(paste("Standard Error=",round(MC_result$SE,2)))
print(paste("Time Taken for Simulation=",round(MC_result$time,2)))
# normalMCput<-Monte_Carlo_Pricing('P',S0,K,T,sigma,div,r,num.steps,num.sim)




```
## B.
Implement a Monte Carlo scheme for European call and put options us-
ing the antithetic variates method (see section 4.3 of the textbook), the
delta{based control variate (section 4.5 of the textbook) with 1 = ????1, and
the combined antithetic variates with delta-based control variate method.
Report the values obtained in four columns: Monte Carlo (MC), MC with
Antithetic Variates, MC with Delta-based Control Variate, and MC with
both Antithetic Variates and Delta-based Control Variate. Report the esti-
mated option values, the corresponding standard deviations, as well as the
time it takes to obtain each result. Write a paragraph comparing the results
you obtained. Discuss the methods implemented.

```{r}

num.sim2<-1000
num.steps2<-300
montecarloAthithetic<-function(PutCall,S0,K,T,sigma,div,r,N,M){
  ## Precompute constants
  dt = T/N
  nudt = (r-div-0.5*sigma^2)*dt
  sigmasdt = sigma*sqrt(dt)
  lns = log(S0)
  
  sum_CT = 0
  sum_CT2 = 0
  start.time = proc.time()
  
  for (j in 1:M){
    lnSt1 = lns
    lnSt2 = lns
    for ( i in 1:N){
      w = rnorm(N)
      lnSt1 = lnSt1 + nudt + sigmasdt*w
      lnSt2 = lnSt2 + nudt + sigmasdt*(-w)
    }
    St1 = exp(lnSt1[N])
    St2 = exp(lnSt2[N])
    optionValue = ifelse(PutCall=='C', 0.5*(pmax(0,St1-K)+pmax(0,St2-K)), 0.5*(pmax(0,K-St1)+pmax(0,K-St2)))
    sum_CT = sum_CT+optionValue
    sum_CT2 = sum_CT2+optionValue*optionValue
  }
  option_value = sum_CT/M*exp(-r*T)
  SD = sqrt((sum_CT2 - sum_CT*sum_CT/M)*exp(-2*r*T)/(M-1))
  SE = SD/sqrt(M)
  end.time = proc.time()
  
  timetaken = end.time - start.time
  list(option_value=option_value, SE=SE, time=timetaken[3])
}

athitheticMC_Call<-montecarloAthithetic('C',S0,K,T,sigma,div,r,num.steps2,num.sim2)
athitheticMC_Put<-montecarloAthithetic('P',S0,K,T,sigma,div,r,num.steps2,num.sim2)

black_scholes_delta<-function(PutCall,S0,t,K,T,r,sigma,div){
  d1<-(log(S0/K)+(r-div+sigma^2/2)*(T-t))/(sigma*sqrt(T-t))
  d2<-d1-sigma*sqrt(T-t)
  value = ifelse(PutCall=='C',exp(-div*(T-t))*pnorm(d1),exp(-div*(T-t))*(pnorm(d1)-1))
  return(value)
}

montecarloDelta<-function(PutCall,S0,K,T,sigma,div,r,beta1,N,M){
  ## Precompute constants
  dt = T/N
  nudt = (r-div-0.5*sigma^2)*dt
  sigmasdt = sigma*sqrt(dt)
  erddt = exp((r-div)*dt)
  
  sum_CT = 0
  sum_CT2 = 0
  start.time = proc.time()
  
  for (j in 1:M){
    St = S0
    cv = 0
    for ( i in 1:N){
      t = (i-1)*dt
      delta = black_scholes_delta(PutCall,St,t,K,T,r,sigma,div)
      w = rnorm(N)
      Stn = St*exp(nudt + sigmasdt*w)
      cv = cv + delta*(Stn-St*erddt)
      St = Stn
    }
    
    optionValue = ifelse(PutCall=='C', (pmax(0,St-K) + beta1*cv), (pmax(0,K-St) + beta1*cv))
    sum_CT = sum_CT+optionValue
    sum_CT2 = sum_CT2+optionValue*optionValue
  }
  option_value = sum_CT/M*exp(-r*T)
  SD = sqrt((sum_CT2 - sum_CT*sum_CT/M)*exp(-2*r*T)/(M-1))
  SE = SD/sqrt(M)
  end.time = proc.time()
  
  timetaken = end.time - start.time
  list(option_value=option_value, SE=SE, time=timetaken[3])
}
beta1<-(-1)  ##given
deltaMC_Call<-montecarloDelta('C',S0,K,T,sigma,div,r,beta1,num.steps2,num.sim2)
deltaMC_Put<-montecarloDelta('P',S0,K,T,sigma,div,r,beta1,num.steps2,num.sim2)


montecarloAthiandDelta<-function(PutCall,S0,K,T,sigma,div,r,beta1,N,M){
  ## Precompute constants
  dt = T/N
  nudt = (r-div-0.5*sigma^2)*dt
  sigmasdt = sigma*sqrt(dt)
  erddt = exp((r-div)*dt)
  
  sum_CT = 0
  sum_CT2 = 0
  start.time = proc.time()
  
  for (j in 1:M){
    St1 = S0
    St2 = S0
    cv1 = 0
    cv2 = 0
    
    for ( i in 1:N){
      t = (i-1)*dt
      delta1 = black_scholes_delta(PutCall,St1,t,K,T,r,sigma,div)
      delta2 = black_scholes_delta(PutCall,St2,t,K,T,r,sigma,div)
      w = rnorm(N)
      Stn1 = St1*exp(nudt + sigmasdt*w)
      Stn2 = St2*exp(nudt + sigmasdt*(-w))
      cv1 = cv1 + delta1*(Stn1-St1*erddt)
      cv2 = cv2 + delta2*(Stn2-St2*erddt)
      St1 = Stn1
      St2 = Stn2
    }
    optionValue = ifelse(PutCall=='C', 0.5*(pmax(0,St1-K) + beta1*cv1 + pmax(0,St2-K) + beta1*cv2), 0.5*(pmax(0,K-St1) + beta1*cv1 + pmax(0,K-St2) + beta1*cv2))
    sum_CT = sum_CT+optionValue
    sum_CT2 = sum_CT2+optionValue*optionValue
  }
  option_value = sum_CT/M*exp(-r*T)
  SD = sqrt((sum_CT2 - sum_CT*sum_CT/M)*exp(-2*r*T)/(M-1))
  SE = SD/sqrt(M)
  end.time = proc.time()
  
  timetaken = end.time - start.time
  list(option_value=option_value, SE=SE, time=timetaken[3])
}
AthiandDeltaMC_Call<-montecarloAthiandDelta('C',S0,K,T,sigma,div,r,beta1,num.steps2,num.sim2)
AthiandDeltaMC_Put<-montecarloAthiandDelta('P',S0,K,T,sigma,div,r,beta1,num.steps2,num.sim2)
```



```{r}
methods=c("MC", "MC-Antithetic Variates", "MC-Delta","MC-AntitheticDelta")
df=data.frame()
df=rbind(df,MC_result,athitheticMC_Call,deltaMC_Call,AthiandDeltaMC_Call)
df$Methods=methods
print(df)

```
