---
title: "Question2"
author: "Saeed Rahman"
date: "May 2, 2017"
output: html_document
---


## A) Heston Monte Carlo
Function to Implement the Heston Monte Carlo for different models in Euler Discretization
```{r}
HestonMC<-function(case,nIterations,nSteps)
{
    
  S_0=100.0
  K = 100.0
  r = 0.0319
  v_0 = 0.010201
  T = 1.00
  
  dt = T/nSteps
  
  rho = -0.7
  kappa = 6.21
  theta = 0.019
  xi = 0.61
  
  rmse=0
  payoff_sum=0
  sumbias=0
  
  ptm <- proc.time()
  for( j in 1:nIterations) {
    W1=W2=0
    vol_path=v_0
    spot_path=S_0
    vmax=0
    for( i in 1:nSteps){
      switch(case,
        Full={# Full truncation
          vmax= max(vol_path,0)
          vol_path=vol_path+kappa*dt*(theta-vmax)+xi*sqrt(vmax*dt)*W1
          spot_path=spot_path* exp((r - 0.5*vmax)*dt +sqrt(vmax*dt)*W2)
        },
        Partial={# Partial truncation
          vmax= max(vol_path,0)
          vol_path=vol_path+kappa*dt*(theta-vol_path)+xi*sqrt(vmax*dt)*W1
          spot_path=spot_path* exp((r - 0.5*vmax)*dt +sqrt(vmax*dt)*W2)
        },
        HM={# Higham and Mao
          vmax= abs(vol_path)
          vol_path=vol_path+kappa*dt*(theta-vol_path)+xi*sqrt(vmax*dt)*W1
          spot_path=spot_path* exp((r - 0.5*vmax)*dt +sqrt(vmax*dt)*W2)
        },
        Refl={# Reflection
          vmax= abs(vol_path)
          vol_path=abs(vol_path)+kappa*dt*(theta-vmax)+xi*sqrt(vmax*dt)*W1
          spot_path=spot_path* exp((r - 0.5*vmax)*dt +sqrt(vmax*dt)*W2)
        },
        Abso={# Absorption
          vmax= max(vol_path,0)
          vol_path=vmax+kappa*dt*(theta-vmax)+xi*sqrt(vmax*dt)*W1
          spot_path=spot_path* exp((r - 0.5*vmax)*dt +sqrt(vmax*dt)*W2)
        }
        )
      
      X1 = rnorm(1,0,1) 
      X2 = rnorm(1,0,1) 
      W1=X1
      W2= rho*X1+sqrt(1-rho^2)*X2
    }
    payoff=max(spot_path-K,0)
    payoff_sum=payoff_sum+payoff
    rmse=rmse+(payoff-6.8061)^2
    sumbias=sumbias+(max(spot_path-K,0)-6.8061)
    # print(max(spot_path-K,0))
  }
  option_price=(payoff_sum/nIterations)*exp(-r*dt)
  time=as.double((proc.time() - ptm)[3])
  rmse=sqrt(rmse)/nIterations
  sumbias=sumbias/nIterations
  
  list(price=option_price,time_taken=time,rmse=rmse,bias=sumbias)
  # print((proc.time() - ptm)[3])
  # print(option_price)
}
```
Completed Table
```{r}
full=    HestonMC(nIterations = 10000,nSteps = 100,case = "Full")
partial= HestonMC(nIterations = 10000,nSteps = 100,case = "Partial")
hm=      HestonMC(nIterations = 10000,nSteps = 100,case = "HM")
refl=    HestonMC(nIterations = 10000,nSteps = 100,case = "Refl")
abso=    HestonMC(nIterations = 10000,nSteps = 100,case = "Abso")

methods=c("Full","Partial","Higham and Mao","Reflection","Absorption")
df=data.frame()
df=rbind(df,full,partial,hm,refl,abso)
df$Methods=methods
print(df)

```


## 2) Bonus: Heston Analytical
```{r setup, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(knitr)
read_chunk('HestonPriceExample.R')
```

```{r Heston_Analytical}
```


We can clearly see that the analytical price is faster.
Among the different monte carlo methods, we can see that accuracy and computation times are inversly related and looking at the above result Higham and Mao gives the fastest result while the accurate price is from the Partial truncation method.
