---
title: "Question2"
author: "Saeed Rahman"
date: "March 26, 2017"
output: html_document
---
## A)
Download Option prices (you can use the Bloomberg Terminal, Yahoo! Fi-
nance, etc.) for an equity, for 3 different maturities (1 month, 2 months,
and 3 months) and 10 strike prices. Use the same method from Homework
1 to calculate the implied volatility. Set the current short-term interest rate
equal to 0.75%.
```{r, message=FALSE, warning=FALSE}
# Retrieving Data and Calculating the Implied Volatility 
  # Calculation for Implied Volatility
    # Black sholes merton pricing function ----
library(quantmod)
BSM<-function(S, K, t, r, sigma,type){
  d1 <- (log(S/K)+(r+sigma^2/2)*t)/(sigma*sqrt(t))
  d2 <- d1 - sigma * sqrt(t)
  if (type == "c")
    result <- S*pnorm(d1) - K*exp(-r*t)*pnorm(d2)
  if (type == "p")
    result <- K*exp(-r*t) * pnorm(-d2) - S*pnorm(-d1)
  return(result)
}
    # Secant function ----
Secant <- function(S, K, t, r, type, option_price
                   , x0=0.1, x1=3, tolerance=1e-07, max.iter=10000){
  x1=3
  theta=.00001
  fun.x1=BSM(S=S,K=K,t=t,r=r,sigma=x1,type=type)-option_price
  count=1
  start.time <- Sys.time()
  while(abs(fun.x1) > tolerance && count<max.iter) {
    x2=x1-theta
    fun.x1=BSM(S=S,K=K,t=t,r=r,sigma=x1,type=type)-option_price
    fun.x2=BSM(S=S,K=K,t=t,r=r,sigma=x2,type=type)-option_price
    x1 <- x1- fun.x1/((fun.x1-fun.x2)/theta)   
    count <-count+1
  }
  end.time <- Sys.time()
  time.taken <- end.time - start.time
  if(x2<0 || count>=max.iter)
    return(list(NA,time.taken,count))
  else
    return(list(x2,time.taken,count))
}

    #Calculating IV on the option chain using secant method
ImpliedVol_Secant<-function(symbol="AAPL",option_chain,rate=.75/100){
  symbol="AAPL"
  stock_df<-as.data.frame(getSymbols(symbol,from = as.Date("2017-01-01"),to=as.Date("2017-02-19"), env = NULL))
  
  iv <- {}
  original_iv <-{}
  optionName <-{}
  strike <-{}
  days_till_expiry <-{}
  time.taken <- 0
  iterations <- 0
  type <-{}
  bid<-{}
  ask<-{}
  for (i in 1:nrow(option_chain)) 
  {
    try({
      #Myoldmethod----
      secant <- Secant(
        S = as.numeric(tail(stock_df,1)[6]),
        K = as.numeric(option_chain[i,"Strike"]),
        t = as.numeric(option_chain[i,"days_till_expiry"])/252,
        r = rate,
        type = ifelse((option_chain[i,"Type"]=="Call"), "c", "p"),
        option_price = as.numeric(option_chain[i,"premium"]))
      
      iv <- append(iv,as.numeric(secant[1]))
      
      if(!is.na(secant[1])){
        time.taken <- as.numeric(secant[2])+time.taken
        iterations <- as.numeric(secant[3])+iterations
      }
      type <- append(type,as.character(option_chain[i,"Type"]))
      
      strike<-append(strike,as.numeric(option_chain[i,"Strike"]))
      
      optionName <- append(optionName,paste(option_chain[i,"Strike"],"-",
                                            option_chain[i,"Type"],"Expiring On:",
                                            option_chain[i,"Expiry"]))
      days_till_expiry <- append(days_till_expiry,as.numeric(option_chain[i,"days_till_expiry"]))
      
      bid<-append(bid,as.numeric(option_chain[i,"Bid"]))
      ask<-append(ask,as.numeric(option_chain[i,"Ask"]))
    })
  }
  option_chain_df <- data.frame(days_till_expiry,type,optionName,iv,strike,bid,ask)
  names(option_chain_df)<-c("Days_till_Expiry","Type","Specification","Implied_Volatility","Strike","Bid","Ask")
  time.taken <- time.taken/as.numeric(colSums(!is.na(option_chain_df))[3])
  iterations <- iterations/as.numeric(colSums(!is.na(option_chain_df))[3])
  list(option_chain_df,time.taken,iterations)
}    
  # Retrieving data from CSV and formating data ----
option_chain_call_csv <- read.csv(file="call.csv",header=TRUE, sep=",")
option_chain_put_csv <-read.csv(file="put.csv",header=TRUE, sep=",")
#Call
option_chain_call_csv$days_till_expiry <- as.Date(option_chain_call_csv$Expiry,"%m/%d/%Y")-as.Date("2017-02-19")
#Put
option_chain_put_csv$days_till_expiry <- as.Date(option_chain_put_csv$Expiry,"%Y/%m/%d")-as.Date("2017-02-19")
#Calculating the days till expiry
option_chain_call_csv$premium<-(option_chain_call_csv$Bid+option_chain_call_csv$Ask)/2

option_chain_put_csv$premium<-(option_chain_put_csv$Bid+option_chain_put_csv$Ask)/2

df=ImpliedVol_Secant(option_chain = option_chain_call_csv)
df=as.data.frame(df[1])
options.df_call=df[complete.cases(df$Implied_Volatility),]

df=ImpliedVol_Secant(option_chain = option_chain_put_csv)
df=as.data.frame(df[1])
options.df_put=df[complete.cases(df$Implied_Volatility),]
```
```{r}
head(options.df_call)
head(options.df_put)
```

## B)
Use the Explicit, Implicit, and Crank-Nicolson Finite Difference schemes
implemented in Problem 1 to price European Call and Put options. Use
the calculated implied volatility to obtain a space dimension ∆x that insures
stability and convergence with an error magnitude of no greater than 0.001.
## C)
For the European style options above (puts and calls) calculate the corre-
sponding Delta, Gamma, Theta, and Vega using the Explicit finite difference
method.
## D)
Create a table with the following columns: time to maturity T , strike price
K, type of the option (Call or Put), ask price A, bid price B, market price
C M = (A + B)/2, implied volatility from the BSM model σ imp , option price
calculated with EFD, IFD, and CNFD. Plot on the same graph A, B, C M ,
and the 3 option prices obtained with finite difference schemes, as a function
of K and T . What can you observe?

## Question B to D done in the Below set of Code

### FD Functions
#### Explicit
```{r, warning=FALSE}
# Explicit Implementation----
Explicit<- function(isCall, K, Tm,S0, r, sig, N, Nj=0, div, dx,returnGreeks=FALSE){
  # Finite Difference Method: i times, 2*i+1 final nodes
  # Precompute constants ----
  if(Nj!=0)
    returnGreeks=FALSE
  dt = Tm/N
  nu = r - div - 0.5 * sig^2
  edx = exp(dx)
  # got the constants formulas from clewlow 3.18, 3.19, 3.20
  pu = 0.5 * dt * ( (sig/dx)^2 + nu/dx )
  pm = 1.0 - dt *   (sig/dx)^2 - r*dt 
  pd = 0.5 * dt * ( (sig/dx)^2 - nu/dx)
  firstRow = 1
  firstCol = 1
  cp = ifelse(isCall, 1, -1)
  if(Nj!=0){
    r = nRows = lastRow = 2*Nj+1
    middleRow = Nj+1
    nCols = lastCol = N+1
    # Intialize asset prices  ----
    V = S = matrix(0, nrow=nRows, ncol=nCols)
    S[middleRow, firstCol] = S0
    S[lastRow,lastCol]= S0*exp(-Nj*dx)
    for(j in (lastRow-1):1){
      S[j,lastCol] = S[j+1,lastCol] * edx
    }
  }
  else{
    Nj=N
    r = nRows = lastRow = 2*Nj+1
    middleRow = s = nCols = lastCol = N+1
    V = S = matrix(0, nrow=nRows, ncol=nCols)
    # Intialize asset prices  ----
    S[middleRow, firstCol] = S0
    for (i in 1:(nCols-1)) {
      for(j in (middleRow-i+1):(middleRow+i-1)) {
        S[j-1, i+1] = S[j, i] * exp(dx)
        S[j ,  i+1] = S[j, i] 
        S[j+1, i+1] = S[j, i] * exp(-dx)
      }
    }
  }
  # Intialize option values at maturity ----
  for (j in 1:lastRow) {
    V[j, lastCol] = max( 0, cp * (S[j, lastCol]-K))
  }
  # Step backwards through the tree ----
  for (i in N:1) {
    for(j in (middleRow+Nj-1):(middleRow-Nj+1)) {
      # This inner for loop is only stepping through the 2 to rowsize-1, to avoid the boundaries
      V[j, i] = pu*V[j-1,i+1] + pm*V[j, i+1] + pd*V[j+1,i+1]
    }
    # Boundary Conditions ----
    stockTerm = ifelse(isCall, S[1, lastCol]-S[2,lastCol],        S[nRows-1,lastCol]-S[nRows,lastCol])
    # The last row contains the discounted value of V[lastRow, lastCol] and since 
    # this is zero for Call, we adopt the below method
    V[lastRow,  i] = V[lastRow-1,  i] + ifelse(isCall, 0, stockTerm)
    # Doing interpolation for Filling the first rows of each column
    V[firstRow, i] = V[firstRow+1, i] + ifelse(isCall, stockTerm, 0)
  }
  # Compute the Greeks ----
  if(returnGreeks && isCall){
    delta = (V[middleRow-1,firstCol+1]-V[middleRow+1,firstCol+1])/
      (S[middleRow-1,firstCol+1]-S[middleRow+1,firstCol+1])
    delta1 =(V[middleRow-1,firstCol+1]-V[middleRow,firstCol+1])/
      (S[middleRow-1,firstCol+1]-S[middleRow,firstCol+1])
    delta2 =(V[middleRow,firstCol+1]-V[middleRow+1,firstCol+1])/
      (S[middleRow,firstCol+1]-S[middleRow+1,firstCol+1])
    gamma = 2*(delta1-delta2)/((S[middleRow-1,firstCol+1]-S[middleRow+1,firstCol+1]))
    theta =((V[middleRow,firstCol+1]-V[middleRow,firstCol])/dt)/252
    return(list(Price=V[middleRow,firstCol],Delta=delta,Gamma=gamma,Theta=theta))
  }
  # Return the price ----
  return(V[middleRow,firstCol])
}
```
#### Implicit
```{r, warning=FALSE}
# Implicit Implementation----
Implicit = function(isCall, K, Tm,S0, r, sig, N, div, dx, Nj=0){
# Implicit Finite Difference Method: i times, 2*i+1 final nodes
# Precompute constants ----
dt = Tm/N
nu = r - div - 0.5 * sig^2
edx = exp(dx)
# got the constants formulas from clewlow 3.33,3.34,3.35
pu = -0.5 * dt * ( (sig/dx)^2 + nu/dx )
pm =  1.0 + dt *   (sig/dx)^2 + r*dt 
pd = -0.5 * dt * ( (sig/dx)^2 - nu/dx)
firstRow = 1
if(Nj!=0){
  r = nRows = lastRow = 2*Nj+1
  middleRow = Nj+1
  nCols = lastCol = N+1
}
else{
  Nj=N
  r = nRows = lastRow = 2*Nj+1
  middleRow = s = nCols = lastCol = N+1
}
firstCol = 1
cp = ifelse(isCall, 1, -1)

# Intialize asset price, derivative price, primed probabilities  ----
pp=pmp=V = S = matrix(0, nrow=nRows, ncol=nCols)
S[middleRow, firstCol] = S0
S[lastRow,lastCol]= S0*exp(-Nj*dx)
for(j in (lastRow-1):1){
  S[j,lastCol] = S[j+1,lastCol] * edx
}
# Intialize option values at maturity ----
for (j in firstRow:lastRow) {
  V[j, lastCol] = max( 0, cp * (S[j, lastCol]-K))
}
# Compute Derivative Boundary Conditions ----
# From equation 3.38 and 3.39 in Clewlow
if(isCall){ 
  lambdaU =(S[1, lastCol] - S[2, lastCol])
  lambdaL = 0
}else{ #clewlows way
  lambdaU = 0
  lambdaL = -1 * (S[lastRow-1, lastCol] - S[lastRow,lastCol])
}
# Step backwards through the lattice ----
for (i in (lastCol-1):firstCol) {
  h = solveImplicitTridiagonal(V, pu, pm, pd, lambdaL, lambdaU, i)
  pmp[,i] = h$pmp  # collect the pm prime probabilities
  pp [,i] = h$pp   # collect the p prime probabilities
  V = h$V
  # Apply Early Exercise condition ----
}
# Return the price ----
return(list(Price=V[middleRow,firstCol],Probs=round(c(pu=pu, pm=pm, pd=pd),middleRow)))
}
# Solving the tridiagonal matrix---- 
solveImplicitTridiagonal=function(V, pu, pm, pd, lambdaL, lambdaU, colI)
{
# Initalize values ----
firstRow = 1
secondRow = 2
thirdRow = 3
lastRow = nRows = nrow(V)
lastCol = ncol(V)
# Substitute boundary condition at j = -Nj into j = -Nj+1 ----
pp = pmp = numeric(nRows)
pmp[lastRow-1] = pm + pd
pp[lastRow-1]  = V[lastRow-1, lastCol] + pd*lambdaL

# Eliminate upper diagonal ----
for (j in (lastRow-2):(secondRow)) {
  pmp[j] = pm - pu*pd/pmp[j+1]
  pp[j] = V[j, colI+1] - pp[j+1]*pd/pmp[j+1]
}
# Use boundary conditions at j = Nj and equation at j=Nj-1 ----
V[firstRow, colI] = (pp[secondRow] + pmp[secondRow]*lambdaU)/(pu + pmp[secondRow])
V[secondRow, colI] = V[firstRow,colI] - lambdaU
# Back-substitution ----
for(j in thirdRow:lastRow) {
  V[j, colI] =  (pp[j] -pu*V[j-1, colI])/pmp[j]
}
V[lastRow, colI] = V[lastRow-1, colI] - lambdaL
# Return values ----
list(V=V, pmp=pmp, pp=pp)
}
```
#### CrankNicholson
```{r, warning=FALSE}
#Crank Nicholson Method----
CrankNicholson = function(isAmerican, isCall, K, Tm,S0, r, sig, N, div, dx, Nj=0){
  # Crank Nicholson Finite Difference Method: i times, 2*i+1 final nodes
  # Precompute constants ----
  dt = Tm/N
  nu = r - div - 0.5 * sig^2
  edx = exp(dx)
  pu = -0.25     *dt * ( (sig/dx)^2 + nu/dx )  
  pm =  1.0 + 0.5*dt *   (sig/dx)^2 + 0.5*r*dt 
  pd = -0.25     *dt * ( (sig/dx)^2 - nu/dx)   
  firstRow = 1
  firstCol = 1
  if(Nj!=0){
    r = nRows = lastRow = 2*Nj+1
    middleRow = Nj+1
    nCols = lastCol = N+1
  }
  else{
    Nj=N
    r = nRows = lastRow = 2*Nj+1
    middleRow = s = nCols = lastCol = N+1
  }
  middleRow = nCols = lastCol = Nj+1
  
  cp = ifelse(isCall, 1, -1)
  
  # Intialize asset price, derivative price, primed probabilities  ----
  pp=pmp=V = S = matrix(0, nrow=nRows, ncol=nCols)
  S[middleRow, firstCol] = S0
  S[lastRow,lastCol]= S0*exp(-Nj*dx)
  for(j in (lastRow-1):1){
    S[j,lastCol] = S[j+1,lastCol] * edx
  }
  # Intialize option values at maturity ----
  for (j in firstRow:lastRow) {
    V[j, lastCol] = max( 0, cp * (S[j, lastCol]-K))
  }
  # Compute Derivative Boundary Conditions ----
  if(isCall){
    lambdaU =(S[1, lastCol] - S[2, lastCol])
    lambdaL = 0
  }else{
    lambdaU = 0
    lambdaL = round(-1 * (S[lastRow-1, lastCol] - S[lastRow,lastCol]),2)
  }
  # Step backwards through the lattice ----
  for (i in (lastCol-1):firstCol) {
    h = solveCrankNicholsonTridiagonal(V, pu, pm, pd, lambdaL, lambdaU, i)
    pmp[,i] = round(h$pmp,4)  # collect the pm prime probabilities
    pp [,i] = round(h$pp, 4)  # collect the p prime probabilities
    V = h$V
    # Apply Early Exercise condition for American Options ----
    for(j in lastRow:firstRow) {
      V[j, i] = max(V[j, i], cp * (S[j, lastCol] - K))
    }
  }
  # Return the price ----
  list(Type = paste(ifelse(isCall, "Call", "Put")),Price = V[middleRow,firstCol],
       Probs=round(c(pu=pu, pm=pm, pd=pd), 4), pmp=pmp, pp= pp,
       S=round(S,2), V=round(V,middleRow))
}
#TridiagonalMatrix solver for crank nicholson----
#it was solveICrankNicholsonTridiagonal
solveCrankNicholsonTridiagonal=function(V, pu, pm, pd, lambdaL, lambdaU, colI)
{
  # Initalize values ----
  firstRow = 1
  secondRow = 2
  thirdRow = 3
  lastRow = nRows = nrow(V)
  lastCol = ncol(V)
  # Substitute boundary condition at j = -Nj into j = -Nj+1 ----
  pp = pmp = numeric(nRows)
  pmp[lastRow-1] = pm + pd
  pp[lastRow-1]  = (- pu   *V[lastRow-2, lastCol] 
                    -(pm-2)*V[lastRow-1, lastCol]
                    - pd   *V[lastRow  , lastCol] + pd*lambdaL)
  # Eliminate upper diagonal ----
  for (j in (lastRow-2):(secondRow)) {
    pmp[j] = pm - pu*pd/pmp[j+1]
    pp[j] = ( - pu   *V[j-1, colI+1] 
              -(pm-2) *V[j  , colI+1]
              - pd    *V[j+1, colI+1] 
              -pp[j+1]*pd/pmp[j+1])
  }               
  # Use boundary conditions at j = Nj and equation at j=Nj-1 ----
  V[firstRow, colI] = (pp[secondRow] + pmp[secondRow]*lambdaU)/(pu + pmp[secondRow])
  V[secondRow, colI] = V[firstRow,colI] - lambdaU
  # Back-substitution ----
  for(j in thirdRow:lastRow) {
    V[j, colI] =  (pp[j] -pu*V[j-1, colI])/pmp[j]
  }
  V[lastRow, colI] = V[lastRow-1, colI] - lambdaL
  # Return values ----
  list(V=V, pmp=pmp, pp=pp)
}



```
#### Calculating Price and Greeks of the downloaded option
```{r, warning=FALSE}
# Calculating option price and dx----
PriceCalculator<-function(options.df,show=FALSE){
  market.price <-{}
  explicit.price <-{}
  implicit.price <-{}
  cranknic.price <-{}
  delta_explicit <-{}
  gamma_explicit <-{}
  theta_explicit <-{}
  vega_explicit  <-{}
  stock_df<-as.data.frame(getSymbols("AAPL",from = as.Date("2017-01-01"),
                                     to=as.Date("2017-02-19"), env = NULL))
  # Current Price = 135.72
  # Calculating dx,N for explicit and implicit based on the error ----
  for (i in 1:nrow(options.df)){
    dt=0
    dx=0
    N=3 #Initial arbitary value
    Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252
    sig = as.numeric(options.df[i,"Implied_Volatility"])
    nsd=3
    error=3
    repeat{
      dt=Tm/N
      Nj=(sqrt(N)*nsd)/(2*sqrt(3)) -.5
      dx=(nsd*sig*sqrt(Tm))/(2*Nj+1)
      N=N+1
      if(((dx*dx)+dt)<=error){
        break
      }
    }
    market.price = append(market.price,((as.numeric(options.df[i,"Bid"]))+(as.numeric(options.df[i,"Ask"])))/2)
    # Calculating the option price using Explicit method ----
    explicit = Explicit(isCall = as.logical(options.df[i,"Type"]=="Call"),
                                K  = as.numeric(options.df[i,"Strike"]),
                                Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252,
                                S0 = as.numeric(tail(stock_df,1)[6]),
                               sig = as.numeric(options.df[i,"Implied_Volatility"]),
                                r  = .75/100,
                               div = 0,
                                dx = dx,
                                 N = N,
                      returnGreeks = TRUE)
    explicit.price= append(explicit.price,as.double(explicit['Price']))
    delta_explicit = append(delta_explicit,explicit['Delta'])
    gamma_explicit = append(gamma_explicit,explicit['Gamma'])
    theta_explicit = append(theta_explicit,explicit['Theta'])
    
    sigma=as.numeric(options.df[i,"Implied_Volatility"])
    dsigma = .001* sigma
    Vega = (Explicit( isCall = as.logical(options.df[i,"Type"]=="Call"),
                          K  = as.numeric(options.df[i,"Strike"]),
                          Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252,
                          S0 = as.numeric(tail(stock_df,1)[6]),
                         sig = sigma+dsigma,
                          r  = .75/100,
                         div = 0,
                          dx = dx,
                           N = N) -
          Explicit(isCall = as.logical(options.df[i,"Type"]=="Call"),
                   K  = as.numeric(options.df[i,"Strike"]),
                   Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252,
                   S0 = as.numeric(tail(stock_df,1)[6]),
                   sig = sigma-dsigma,
                   r  = .75/100,
                   div = 0,
                   dx = dx,
                   N = N))/(2*dsigma*100)
    vega_explicit = append(vega_explicit,Vega)
    
    # Calculating the option price using Implicit method ----
    implicit.price <- append(implicit.price,Implicit(
      isCall = as.logical(options.df[i,"Type"]=="Call"),
      K  = as.numeric(options.df[i,"Strike"]),
      Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252,
      S0 = as.numeric(tail(stock_df,1)[6]),
      sig = as.numeric(options.df[i,"Implied_Volatility"]),
      r  = .75/100,
      div = 0,
      dx = dx,
      N = N)$Price)
    
    
    #calculating dx and N using convergence condition for crank nicholson----
    dt=0
    dx=0
    N=3 #Initial arbitary value
    Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252
    sig = as.numeric(options.df[i,"Implied_Volatility"])
    nsd=3
    error=3
    repeat{
      dt=Tm/N
      Nj=(sqrt(N)*nsd)/(2*sqrt(3)) -.5
      dx=(nsd*sig*sqrt(Tm))/(2*Nj+1)
      N=N+1
      if(((dx*dx)+(dt/2))<=error){
        break
      }
    }
    
    # Calculating the option price using Cranknicholson method ----
    cranknic.price <- append(cranknic.price,CrankNicholson(
      isCall = as.logical(options.df[i,"Type"]=="Call"),
      K  = as.numeric(options.df[i,"Strike"]),
      Tm = as.numeric(options.df[i,"Days_till_Expiry"])/252,
      S0 = as.numeric(tail(stock_df,1)[6]),
      sig = as.numeric(options.df[i,"Implied_Volatility"]),
      r  = .75/100,
      div = 0,
      dx = dx,
      N = N)$Price)

  }
  options.df$Market.price <- market.price
  options.df$Explicit.price <- explicit.price
  options.df$Implicit.price <- implicit.price
  options.df$CrankNicholson.price <- cranknic.price
  options.df$Delta.explicit = delta_explicit
  options.df$Gamma.explicit = gamma_explicit
  options.df$Theta.explicit = theta_explicit
  options.df$Vega.explicit  = vega_explicit
  if(show){
    head(options.df)
    return(options.df)
  }
  else
    return(options.df)
}

# price=PriceCalculator(options.df_call,show=TRUE)
option_chain_call <-PriceCalculator(options.df_call)
option_chain_put <-PriceCalculator(options.df_put)
head(option_chain_call)
```
### D)Plot


```{r}

theTitle = expression(paste("Finite Difference Price Plot 2"))
            
xlab1 = expression(paste("Days Till Expiry"))
plot(option_chain_call$Days_till_Expiry, option_chain_call$Explicit, ylim=c(0,50), type="n", xlab="")
title(theTitle, cex=5)
mtext(text=xlab1, side=1, line=2, cex=1.25)
lines(option_chain_call$Days_till_Expiry, option_chain_call$Implicit, lwd=2, col=2)
lines(option_chain_call$Days_till_Expiry, option_chain_call$CrankNicholson, lwd=2, col=3)
lines(option_chain_call$Days_till_Expiry, option_chain_call$Market.price, lwd=2, col=4)
legend(50, 40, legend=c('Explicit','Implicit','CrankNS','Market Price'), col=2:5, lty=1, lwd=3)
```
