---
title: "Assignment 3 - Question1"
author: "Saeed Rahman"
date: "March 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A)
Implement the Explicit Finite Difference method to price both European
Call and Put options.

```{r}

```

## B)
Implement the Implicit Finite Difference method to price European Call
and Put options.

```{r pressure, echo=FALSE}
plot(pressure)
```

## C)
For both the Explicit and Implicit Finite Difference schemes estimate the
numbers ∆t, ∆x as well as the total number N j of points on the space grid
x to obtain a desired error of ε = 0.001.

## D)
Consider S 0 = 100, K = 100, T = 1 year, σ = 25%, r = 6%, δ = 0.03.
Calculate and report the price for European Call and Put using both explicit
and implicit FD methods and the number of steps that you calculated in
the previous point (part c).

## E)
Repeat part (c) of this problem but this time get the empirical number of
iterations. Specifically, obtain the Black Scholes price for the data in (d),
then do an iterative procedure to figure out the ∆x, ∆t, N , and N j to
obtain an accuracy of ε = 0.001.

## F)
Using the parameters from part (d), plot on the same graph the implicit fi-
nite difference probabilities p u , p m , p d as a function of σ, σ ∈ {0.05, 0.1, 0.15, ..., 0.6}.
Write detailed comments on your observations.

## G)
Implement the Crank-Nicolson Finite Difference method and price both
European Call and Put options. Use the same parameters as in part (d)
and the same number of steps in the grid. Put the results of the 3 methods
(EFD, IFD, CNFD) side by side in a table and write your observations.

## H)
Calculate the hedge sensitivities for the European call option using the
Explicit Finite Difference method. You need to calculate Delta, Gamma,
Theta, and Vega.

